Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV statistical mechanics

### 2020-05-21 paper2_JNLS/Referee Reports/Ref1Response/Ref1.tex
NM

\noindent
\\We thank the referee for these positive remarks. \\ \\

@@

\begin{itemize}
\item Thank you for these insightful comments which have helped us to improve the discussion of the Gibbs measure throughout the paper. We have added revisions to several key places of the paper, detailed as follows:
\item First, The referee is correct that a sufficient amount of nonlinear interaction is  required for relaxation to the Gibbs measure. We agree that it is most likely the failure of this condition being met that explains the discrepancy between formula (53) and the smallest-amplitude data in Fig. 8(c). We have inserted these points in the updated manuscript. First, see the passage in Section 3.5 (page 11):
\begin{quotation}
We also point out that a sufficient amount of nonlinear mixing is required for relaxation to the above Gibbs measure. In the linearized system, for example, modes propagate without interacting and hence the initial measure induced by the wave generator remains unchanged.
\end{quotation}
Second, see the passage describing Fig. 8(c) in Section 4.4 (page 22)
\begin{quotation}
The apparently less predictable behavior at small amplitudes can be attributed to the fact that, in order for the system to relax to the Gibbs measure, modes must interact through nonlinearity. For the smallest amplitude experiments, the amount of nonlinearity is apparently too small for  modes to mix sufficiently.
\end{quotation}
%
\item We thank the referee for drawing our attention to the vibrant field of integrable turbulence. We have now included a discussion to compare and contrast our statistical theory with those stemming from integrable turbulence (and included a few key references). In particular, please see the new passage at the end of section 3.4
\begin{quotation}
We comment on some important differences between the KdV and TKdV dynamical systems. Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants [26, 52]. Trajectories of KdV are therefore constrained by an infinite number of conserved quantities, associated with the fact that it is a completely integrable system whose long-time dynamics are dominated by coherent structures known as solitons. The analogues of the additional invariants, however, are not generally conserved in the truncated system; the only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian. Dynamics of TKdV, rather than being dominated by coherent structures, become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though a rigorous proof awaits) [2]. In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence [54, 40, 10, 41], where integrability is exploited to uncover certain statistical features of the system at hand. We remind the reader that KdV is asymptotically valid only for large spatial scales, and hence truncating the small scales does not alter the order of approximation made. In this sense, KdV and TKdV are on equal footing as far as physical realism – a more accurate framework would require either model closure to capture the small scale physics or else direct numerical simulation of the Euler equations. Where differences exist between KdV and TKdV, therefore, it is unclear a priori which model, if either, more closely resembles reality. This situation, of course, underscores the importance of controlled laboratory experiments.
\end{quotation}
\end{itemize}
%\item The referee is correct that equilibrium statistical mechanics and the canonical ensemble is used here, not because it is rigorously justified, but as an effective theory that explains the data.

>>>>

\noindent
\\We thank the referee for this positive summary of our work. \\ \\

@@

\begin{itemize}
\item Thank you for these insightful comments which have helped us to improve the discussion of the Gibbs measure throughout the paper. We have added revisions to several key places of the paper, detailed as follows:
\item First, The referee is correct that a sufficient amount of nonlinear interaction is  required for relaxation to the Gibbs measure. We agree that it is most likely the failure of this condition being met that explains the discrepancy between formula (53) and the smallest-amplitude data in Fig. 8(c). We have inserted these points in the updated manuscript. First, see the passage in Section 3.5 (page 11):
\begin{quotation}
We also point out that a sufficient amount of nonlinear mixing is required for relaxation to the above Gibbs measure. In the linearized system, for example, modes propagate without interacting and hence the initial measure induced by the wave generator remains unchanged.
\end{quotation}
Second, see the passage describing Fig. 8(c) in Section 4.4 (page 22)
\begin{quotation}
The apparently less predictable behavior at small amplitudes can be attributed to the fact that, in order for the system to relax to the Gibbs measure, modes must interact through nonlinearity. For the smallest amplitude experiments, the amount of nonlinearity is apparently too small for  modes to mix sufficiently.
\end{quotation}
%
\item We thank the referee for drawing our attention to the vibrant field of integrable turbulence. We have now included a discussion to compare and contrast our statistical theory with those stemming from integrable turbulence (and included a few key references). In particular, please see the new paragraphs at the end of section 3.4:
\begin{quotation}
We comment on some important differences between the KdV and TKdV dynamical systems. Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants [26, 52]. Trajectories of KdV are therefore constrained by infinitely many conserved quantities, associated with the fact that it is a completely integrable system. The long-time dynamics of KdV are dominated by coherent structures, namely solitons. With finite truncation, however, the additional invariants cease to be conserved. The only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian. Thus, finite truncation breaks integrability. Moreover, although soliton solutions are supported by TKdV, the long-time dynamics are not dictated by these structures. Rather, the finite truncation induces chaotic to weakly turbulent dynamics, as has been observed numerically [2, 31]. Stronger mixing is induced with energy cascades down and up scales.

In this way, the theory developed herein differs from, and could likely complement, recent
advances in the field of integrable turbulence [54, 40, 10, 41], where integrability is exploited to uncover certain statistical features of the system at hand. We remind the reader that KdV is asymptotically valid only for large spatial scales, and hence truncating the small scales does not alter the order of approximation made. In this sense, KdV and TKdV are on equal footing as far as physical realism – a more accurate framework would require either model closure to capture the small scale physics or else direct numerical simulation of the Euler equations. Where differences exist between KdV and TKdV, therefore, it is unclear a priori which model, if either, more closely resembles reality. This situation, of course, underscores the importance of controlled laboratory experiments.
\end{quotation}
\end{itemize}
%\item The referee is correct that equilibrium statistical mechanics and the canonical ensemble is used here, not because it is rigorously justified, but as an effective theory that explains the data.

### 2020-05-21 paper2_JNLS/Referee Reports/Ref2Response/Ref2.tex
NM

\item Finally, we would like to point out that, since the KdV equation is asymptotically valid only for large spatial scales, increasing $\Lambda$ does not necessarily improve the physical accuracy of the model past a certain point. These small scales lie outside the range of validity of long-wave theory and a more accurate physical description would require either direct numerical simulation of the Euler equations or else a model closure that captures the small-scale physics. We have included this point in the updated manuscript in Section 3.4:
\begin{quotation}
We remind the reader that KdV is asymptotically valid only for large spatial scales, and hence truncating the small scales does not alter the order of approximation made. In this sense, KdV and TKdV are on equal footing as far as physical realism – a more accurate framework would require either model closure to capture the small scale physics or else direct numerical simulation of the Euler equations. Where differences exist between KdV and TKdV, therefore, it is unclear a priori which model, if either, more closely resembles reality. This situation, of course, underscores the importance of controlled laboratory experiments.
\end{quotation}
\end{itemize}

>>>>

\item Finally, we would like to point out that, since the KdV equation is asymptotically valid only for large spatial scales, increasing $\Lambda$ does not necessarily improve the physical accuracy of the model past a certain point. These small scales lie outside the range of validity of long-wave theory and a more accurate physical description would require either direct numerical simulation of the Euler equations or else a model closure that captures the small-scale physics. We have included this point in the updated manuscript in Section 3.4:
\begin{quotation}
We comment on some important differences between the KdV and TKdV dynamical systems. Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants [26, 52]. Trajectories of KdV are therefore constrained by infinitely many conserved quantities, associated with the fact that it is a completely integrable system. The long-time dynamics of KdV are dominated by coherent structures, namely solitons. With finite truncation, however, the additional invariants cease to be conserved. The only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian. Thus, finite truncation breaks integrability. Moreover, although soliton solutions are supported by TKdV, the long-time dynamics are not dictated by these structures. Rather, the finite truncation induces chaotic to weakly turbulent dynamics, as has been observed numerically [2, 31]. Stronger mixing is induced
with energy cascades down and up scales.
\end{quotation}
\end{itemize}

### 2020-05-21 paper2_JNLS/paper.tex
NM

\author{
M.~N.~J.~Moore\thanks{United States Naval Academy}, 
C.~Tyler Bolles\thanks{University of Michigan},
Andrew J.~Majda\thanks{Courant Institute of Mathematical Sciences}, 
Di Qi\footnotemark[3] }
\maketitle

@@

We remark that, in defining the Hamiltonian, we have chosen the sign convention of Lax (1975) \cite{lax1975periodic}. More recent work of Bajars {\it et al.}~(2013) \cite{bajars2013weakly} and \maetal \cite{majda2019} use a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite. Clearly, these sign differences cancel in \eqref{HamStruct} and thus the two conventions are completely equivalent. Using the second convention, \maetal found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}. We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most standard form of statistical mechanics.

@@

Remarkably, the TKdV system \eqref{TKdV} retains the Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019}. The piecewise defined Hamiltonian for TKdV is given by
\begin{equation}
\label{TruncHamiltonian}
\HLupdn = C_2 \dratupdn^{1/2} \, \Htwo[\uL] - C_3 \dratupdn^{-3/2} \, \Hthree[\uL]
\end{equation}
where $\Hthree$ and $\Htwo$ are defined exactly as before \eqref{H3H2}, but now are simply applied to the projected variable $\uL = \Proj u$.
Then TKdV \eqref{TKdV} can be expressed as
\begin{align}
\partial_t {\uL} = \partial_x \Proj \, \vard{\HLupdn}{\uL}
\end{align}
where the truncated symplectic operator is $\SympL = \partial_x \Proj$.
% Note: Bajars 2013 has some important details on the Hamiltonian on the truncated system.

The momentum and energy defined in \eqref{MomEn} remain invariants of TKdV, with the same normalized values $\Mo[\uL] = 0$ and $\En[\uL] = 1$. Note that Parseval's identity implies
\begin{equation}
\En[\uL] = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2 = 1
\end{equation}
Thus, the dynamics of interest are confined to the unit hypersphere, $\En = 1$ in $\CC^{\Lambda}$.
%In summary, the TKdV system possesses three important invariants: momentum, energy, and Hamiltonian.

\new{We comment on some important differences between the KdV and TKdV dynamical systems. Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}. Trajectories of KdV are therefore constrained by an infinite number of conserved quantities, associated with the fact that it is a completely integrable system whose long-time dynamics are dominated by coherent structures known as solitons. The analogues of the additional invariants, however, are {\em not} generally conserved in the truncated system; the only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian. Dynamics of TKdV, rather than being dominated by coherent structures, become chaotic in long time and have been numerically observed to satisfy weak ergodicity (though a rigorous proof awaits) \cite{bajars2013weakly}. In this way, the theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence \cite{zakharov2009turbulence, randoux2014intermittency, costa2014soliton, randoux2016nonlinear}, where  integrability is exploited to uncover certain statistical features of the system at hand. We remind the reader that KdV is asymptotically valid only for large spatial scales, and hence truncating the small scales does not alter the order of approximation made. In this sense, KdV and TKdV are on equal footing as far as physical realism -- a more accurate framework would require either model closure to capture the small scale physics or else direct numerical simulation of the Euler equations. Where differences exist between KdV and TKdV, therefore, it is unclear {\em a priori} which model, if either, more closely resembles reality. This situation, of course, underscores the importance of controlled laboratory experiments.
}

@@

	Now, from the perspective of statistical mechanics, consider the communication between the statistical ensembles, $\Gupdn$, upstream and downstream of the ADC. These two systems are in contact at the ADC, and so the upstream state with distribution $\Gup$ can be regarded as a thermal reservoir that influences the downstream distribution $\Gdn$. In fact, the above deterministic  condition directly leads to a simple description for the link. The quantity of interest is the outgoing Hamiltonian $\Hdn$, since its value just upstream of the ADC, $t=0^{-}$, is set by the incoming dynamics, and then this particular value is conserved thereafter in the outgoing dynamics. Since $u$ matches at the ADC, so must $\Hdn$, and, in fact, this matching holds for every individual trajectory. Recall that $\Hdn(t)$ is not conserved in the upstream dynamics and so its value varies for $t < 0$. However, appealing to weak Ergodicity, the ensemble mean $\meanup{\Hdn(t)}$ is expected to be independent of time. In particular, the value $\meanup{\Hdn \vert_{t=0^{-}} }$ is the same as the bare ensemble mean $\meanup{\Hdn}$. Afterwards, the particular value $\Hdn \vert_{t=0^{-}}$ is conserved in the downstream dynamics, on a trajectory-by-trajectory basis and for all $t>0$. Thus, the downstream measure $\Gdn$ must recover the same ensemble mean $\meanup{\Hdn \vert_{t=0^{-}} } = \meanup{\Hdn}$, producing the simple condition
\begin{align}
\label{statmatch}
\meanup{\Hdn} = \meandn{\Hdn}
\qquad \mbox{\em Statistical matching condition}
\end{align}

>>>>

\author{
Nicholas~J.~Moore\thanks{United States Naval Academy}, 
C.~Tyler Bolles\thanks{University of Michigan},
Andrew J.~Majda\thanks{Courant Institute of Mathematical Sciences}, 
Di Qi\footnotemark[3] }
\maketitle

@@

We remark that, in defining the Hamiltonian, we have chosen the sign convention of Lax (1975) \cite{lax1975periodic}. More recent work of Bajars {\it et al.}~(2013) \cite{bajars2013weakly} and \maetal \cite{majda2019} use a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite. Clearly, these sign differences cancel in \eqref{HamStruct} and thus the two conventions are completely equivalent. Using the second convention, \maetal found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}. We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most conventional form of statistical mechanics.

@@

Remarkably, the TKdV system \eqref{TKdV} retains the Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019}. The piecewise defined Hamiltonian for TKdV is given by
\begin{equation}
\label{TruncHamiltonian}
\HLupdn = C_2 \dratupdn^{1/2} \, \Htwo[\uL] - C_3 \dratupdn^{-3/2} \, \Hthree[\uL]
\end{equation}
where $\Hthree$ and $\Htwo$ are defined exactly as before \eqref{H3H2}, but now are simply applied to the projected variable $\uL = \Proj u$.
Then TKdV \eqref{TKdV} can be expressed as
\begin{align}
\partial_t {\uL} = \partial_x \Proj \, \vard{\HLupdn}{\uL}
\end{align}
where the truncated symplectic operator is $\SympL = \partial_x \Proj$.

The momentum and energy defined in \eqref{MomEn} remain invariants of TKdV, with the same normalized values $\Mo[\uL] = 0$ and $\En[\uL] = 1$. Note that Parseval's identity implies
\begin{equation}
\En[\uL] = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2 = 1
\end{equation}
Thus, the dynamics of interest are confined to the unit hypersphere, $\En = 1$ in $\CC^{\Lambda}$.

\new{We comment on some important differences between the KdV and TKdV dynamical systems. Besides momentum, energy, and Hamiltonian, KdV possesses in infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}. Trajectories of KdV are therefore constrained by infinitely many conserved quantities, associated with the fact that it is a completely integrable system. The long-time dynamics of KdV are dominated by coherent structures, namely solitons. With finite truncation, however, the additional invariants cease to be conserved. The only known invariants of TKdV are the three mentioned already: momentum, energy, and Hamiltonian. Thus, finite truncation breaks integrability. Moreover, although soliton solutions are supported by TKdV, the long-time dynamics are not dictated by these structures. Rather, the finite truncation induces chaotic to weakly turbulent dynamics, as has been observed numerically \cite{bajars2013weakly, majdaqi2019}. Stronger mixing is induced with energy cascades down and up scales.

In this way, the statistical theory developed herein differs from, and could likely complement, recent advances in the field of integrable turbulence \cite{zakharov2009turbulence, randoux2014intermittency, costa2014soliton, randoux2016nonlinear}, where  integrability is exploited to uncover certain statistical features of the system at hand. We remind the reader that KdV is asymptotically valid only for large spatial scales, and hence truncating the small scales does not alter the order of approximation made. In this sense, KdV and TKdV are on equal footing as far as physical realism -- a more accurate framework would require either model closure to capture the small scale physics or else direct numerical simulation of the Euler equations. Where differences exist between KdV and TKdV, therefore, it is unclear {\em a priori} which model, if either, more closely resembles reality. This situation, of course, underscores the importance of controlled laboratory experiments.
}

@@

	Now, from the perspective of statistical mechanics, consider the communication between the statistical ensembles, $\Gupdn$, upstream and downstream of the ADC. These two systems are in contact at the ADC, and so the upstream state with distribution $\Gup$ can be regarded as a thermal reservoir that influences the downstream distribution $\Gdn$. In fact, the above deterministic  condition directly leads to a simple description for the link. The quantity of interest is the outgoing Hamiltonian $\Hdn$, since its value just upstream of the ADC, $t=0^{-}$, is set by the incoming dynamics, and then this particular value is conserved thereafter in the outgoing dynamics. Since $u$ matches at the ADC, so must $\Hdn$, and, in fact, this matching holds for every individual trajectory. Recall that $\Hdn(t)$ is not conserved in the upstream dynamics and so its value varies for $t < 0$. However, appealing to weak ergodicity, the ensemble mean $\meanup{\Hdn(t)}$ is expected to be independent of time. In particular, the value $\meanup{\Hdn \vert_{t=0^{-}} }$ is the same as the bare ensemble mean $\meanup{\Hdn}$. Afterwards, the particular value $\Hdn \vert_{t=0^{-}}$ is conserved in the downstream dynamics, on a trajectory-by-trajectory basis and for all $t>0$. Thus, the downstream measure $\Gdn$ must recover the same ensemble mean $\meanup{\Hdn \vert_{t=0^{-}} } = \meanup{\Hdn}$, producing the simple condition
\begin{align}
\label{statmatch}
\meanup{\Hdn} = \meandn{\Hdn}
\qquad \mbox{\em Statistical matching condition}
\end{align}
